---
title: "GM AI hex maps"
author:
  - name: "Dan Olner"
date: 09/03/2025
format: 
  html:
    output-file: "GM_AI_hexmaps"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
collapse: false
---

```{r setup}
library(tidyverse)
library(ggrepel)
library(sf)
library(tmap)
library(ggridges)
library(plotly)
source('../../functions/helpers.R')
source('../../functions/ad_hoc_functions.R')

options(scipen = 999)

#Set ggplot theme
theme_set(theme_light())

#Get AIOE lookup
soc_3dig_AIOE <- readRDS('../../data/UK_SOC_3digit_linkedAIOE.rds')
#soc_3dig_AIOE <- readRDS('data/UK_SOC_3digit_linkedAIOE.rds')

#Get map data
overlay <- readRDS('../../local/data/AIE4measures_hexoverlay.rds')
sq <- readRDS('../../local/data/sq_forhexoverlay.rds')

# itl2 <- st_read('../../../RegionalEcons_web/data/ITL_geographies/International_Territorial_Level_2_January_2021_UK_BFE_V2_2022_-4735199360818908762/ITL2_JAN_2021_UK_BFE_V2.shp',quiet = TRUE) %>% st_simplify(preserveTopology = T, dTolerance = 100) %>% 
#   filter(qg('manchester',ITL221NM))
# 
# lad <- st_read("~/Dropbox/MapPolygons/UK/2024/Local_Authority_Districts_May_2024_Boundaries_UK_BFC/LAD_MAY_2024_UK_BFC.shp",quiet = TRUE) %>% st_simplify(preserveTopology = T, dTolerance = 100)

#Blue peter'd those up just to GM
itl2 <- readRDS('../../local/data/itl2_for_GM.rds')
lad <- readRDS('../../local/data/las_for_GM.rds')

```


## Bivariate hexmaps for AIIE vs "aug \> replace"

This and the next few plots are **bivariate maps** - each axis has three values of 'low / medium / high', making **nine possible combinations** of the two of them.

These two maps:

-   Puts 'probability job is more augmentable than replaceable' **(aug\>repl)** along one axis, with Felten's **AIIE** exposure index on the other.
-   The first is for **firms with 1 to 3 employees**, the second for **firms with 4+ employees**.
-   Data for firms is summarised per 1000m hex (averages weighted by employee count and across any of the , as above).
-   Hexes are only shown if they have a **min of 10 employees in them**.

So for example, in the map of firms with 1 to 3 employees:

-   The darkest hexes in the south of Greater Manchester are high aug\>rep (highest probability of jobs being more augmentable than replaceable) and also high AIIE (they have most AI exposure).
-   There are only a few hexes with the legend's top left clear blue - high exposure but jobs more likely to be replaceable.
-   In both maps, there are large areas of high exposure with the middle aug\>rep probability (jobs likely a mix of both augmentable and replaceable).

### 1 to 3 employees

```{r, fig-hex1_aiie_augmorethanreplace1to3}

employeecountperhexmin = 1

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  # filter(Employees_thisyear > 9) %>% 
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= employeecountperhexmin) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(lad) +
  tm_polygons(fill = '#a6baa8') +
  # tm_polygons(fill = '#93ba98') +
  # tm_polygons(fill = '#5f6e61') +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) +
tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_polygons(
    fill = tm_vars(c("AIIE", "aug>rep"),
    # fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) 
  


```

### 4+ employees

```{r, fig-hex1_aiie_augmorethanreplace4plus}

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  # filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  filter(Employees_thisyear > 4) %>%
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= employeecountperhexmin) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(lad) +
  tm_polygons(fill = '#a6baa8') +
  # tm_polygons(fill = '#93ba98') +
  # tm_polygons(fill = '#5f6e61') +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) +
tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_polygons(
    fill = tm_vars(c("AIIE", "aug>rep"),
    # fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) 
  # tm_shape(itl2) +
  # tm_borders(col = 'black', lwd = 4) 


```

# 'Augmentable' and 'replaceable' each on their own scale

Here, two *separate* indices have been generated that each rank sectors as 'jobs more/less likely to be (1) replaceable and (2) augmentable in this sector compared to other sectors'.

These two indices are made by 'competing' repeated sectors against each other (using the probabilities assigned from the LLM) and putting these pairwise outcomes into a [Bradley Terry](https://en.wikipedia.org/wiki/Bradley%E2%80%93Terry_model){target="_blank"} model to produce a 'sector x more likely to be augmentable than y' ranking. These are then mean-centred (keeping the relative log-likelihoods between sectors intact).

## Where (270 3 digit SIC) sectors are located on the separate aug / replace axes

This first interactive plot just shows those two scales on each axis to illustrate where the method places each sector - **hover for sector name**.

```{r}

bt_trials_wsectors <- readRDS('../../local/data/bt_trials.rds')
#bt_trials_wsectors <- readRDS('local/data/bt_trials.rds')

p <- ggplot(bt_trials_wsectors,aes(x = replace, y = augment, label = SIC_3DIGIT_NAME)) +
  geom_point(size = 3) +
  geom_vline(xintercept = 0, colour = 'darkgrey') +
  geom_hline(yintercept = 0, colour = 'darkgrey') 

ggplotly(p, tooltip = 'label')




```

## Separate augment / replace hexmaps

These maps put 'probability augmentable (high/medium/low)' and 'probability replaceable (high/medium/low)' on separate axes, again meaning there are nine possible values.

As above, there is one map for 'firms with 1 to 3 employees' and a second for 'firms with 4+ employees'.

Some comments on these two maps:

-   There's an obvious stark difference for 1-3 employee vs 4+ employee firms - for 1-3s, there are large areas with lower augmentable and medium to high replacable.
-   Though there's a cluster (matching the one above for aug\>repl vs AIIE) that's low replaceable / medium augmentable (which fits with 'prob aug is more than prob replace' but here we can get a better idea of the actual scale of both).
-   The 4+ map has a lot more hexes in the max category for *both* aug and replace.

### Firms with 1 to 3 employees

```{r, fig-hexmap_augreplace1to3}

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  # filter(Employees_thisyear > 4) %>%
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= employeecountperhexmin) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(lad) +
  tm_polygons(fill = '#a6baa8') +
  # tm_polygons(fill = '#93ba98') +
  # tm_polygons(fill = '#5f6e61') +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) +
tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_polygons(
    # fill = tm_vars(c("AIIE", "aug>rep"),
    fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) 
  # tm_shape(itl2) +
  # tm_borders(col = 'black', lwd = 4) 


```

### Firms with 4+ employees

```{r, fig-hexmap_augreplace4plus}

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  # filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  filter(Employees_thisyear > 4) %>%
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= employeecountperhexmin) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(lad) +
  tm_polygons(fill = '#a6baa8') +
  # tm_polygons(fill = '#93ba98') +
  # tm_polygons(fill = '#5f6e61') +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) +
tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_polygons(
    # fill = tm_vars(c("AIIE", "aug>rep"),
    fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) 
  # # tm_shape(itl2) +
  # tm_borders(col = 'black', lwd = 4) 


```


```{r eval= F}
# Hexmap for interaction, putting here for ease of making
tmap_mode('view')
tmap_mode('plot')

employeecountperhexmin = 1

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  # filter(Employees_thisyear > 9) %>% 
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= employeecountperhexmin) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_basemap(c(StreetMap = "OpenStreetMap")) +
  tm_polygons(fill_alpha = 0.7,
    fill = tm_vars(c("AIIE", "aug>rep"),
    # fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) 



```



