---
title: "GM AI misc plots"
author:
  - name: "Dan Olner"
date: 09/03/2025
format: 
  html:
    output-file: "GM_AI_miscplots"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
collapse: false
---

```{r setup}
library(tidyverse)
library(nomisr)
library(sf)
library(tmap)
library(ggridges)
library(plotly)
source('../../functions/helpers.R')
source('../../functions/ad_hoc_functions.R')

options(scipen = 999)

#Set ggplot theme
theme_set(theme_light())

#Get AIOE lookup
soc_3dig_AIOE <- readRDS('../../data/UK_SOC_3digit_linkedAIOE.rds')

#Get map data
overlay <- readRDS('../../local/data/AIE4measures_hexoverlay.rds')
sq <- readRDS('../../local/data/sq_forhexoverlay.rds')

# itl2 <- st_read('../../../RegionalEcons_web/data/ITL_geographies/International_Territorial_Level_2_January_2021_UK_BFE_V2_2022_-4735199360818908762/ITL2_JAN_2021_UK_BFE_V2.shp',quiet = TRUE) %>% st_simplify(preserveTopology = T, dTolerance = 100) %>% 
#   filter(qg('manchester',ITL221NM))
# 
# lad <- st_read("~/Dropbox/MapPolygons/UK/2024/Local_Authority_Districts_May_2024_Boundaries_UK_BFC/LAD_MAY_2024_UK_BFC.shp",quiet = TRUE) %>% st_simplify(preserveTopology = T, dTolerance = 100)

#Blue peter'd those up just to GM
itl2 <- readRDS('../../local/data/itl2_for_GM.rds')
lad <- readRDS('../../local/data/las_for_GM.rds')

```

# Greater Manchester AI exposure, augmentability and replaceability

The **first section** applies data from [Felten et al](https://sms.onlinelibrary.wiley.com/doi/full/10.1002/smj.3286){target="_blank"} paper "Occupational, industry, and geographic exposure to artificial intelligence: A novel dataset and its potential uses" (2021). Original US data available [here](https://github.com/AIOE-Data/AIOE){target="_blank"}.

The **second section** presents results from a 'jobs replaceable versus augementable' analysis, looking at those separately as well as linking them to Felten's AIIE. 

Both are linked to Companies House firm level data for Greater Manchester.

[To be done: explanation of link process. All code for that is [here](https://github.com/DanOlner/AI_economy/tree/main){target="_blank"}.]

## Links to interactive maps

* [Map 1:](https://danolner.github.io/AI_economy/maps/GM_AI_FeltenAIIE_avweightedbyemployees_CompaniesHouse.html){target="_blank"} Greater Manchester AI Industry Exposure hexmap (opens in new tab). Using Companies House data, all firm sizes.
* [Map 2:](https://danolner.github.io/AI_economy/maps/GM_AI_FeltenAIIE_avweightedbyemployees_CompaniesHouseFirms1to3employees.html){target="_blank"} Same but using **just** firms with 1 to 3 employees.



## AI Industry Exposure (AIIE) applied to UK Companies House data

Felten's AIIE is applied to individual firms in UK Companies House data [method]. This is done for accounts filed July 2024-25. 

Here, these results are **summarised per MSOA** in Greater Manchester (and core cities for comparison) and **broken down into three groups**: firms with 1-3 employees, firms with 4 to 9 employees and 9+ employee firms.

* AIIE on the x axis. Higher values are more exposed firms. Data is normalised, so an AIIE of zero is the mean value and 1 is a single standard deviation.
* An AIIE average (weighted by employee number) is found for each MSOA, broken into the 3 firm size groups.
* Local authorities are ordered top to bottom by average AIIE overall. Greater Manchester boroughs are larger, bolder points; core cities are smaller/fainter.

Three things jump out. 

1. Firms with 1 to 3 employees (green down-pointing triangles) are generally in more exposed sectors than other sized firms (though with overlap) and clearly more exposed than firms with 4-9 employees (red squares).
2. That pattern is stronger in some places (Trafford, Stockport, Bolton) than others
3. The overall AIIE exposure level per borough is mostly consistent within them, for different size firms. So Rochdale, Oldham, Tameside and Wigan are overall lower on the AIIE measure.




```{r}
#| fig-height: 11

#For 'both' see 'all three!'
#Made here:
#https://github.com/DanOlner/AI_economy/blob/7127bf7e022198965d409bd71444767428bb2da4/sector_linkages.R#L876
msoa_both <- readRDS('../../local/data/msoa_3firmsizes_AIIE.rds')

ggplot(msoa_both %>% filter(type != 'other'), 
       aes(
         x = AIIE_weightedbyemployees, 
         y = fct_reorder(localauthority_name,AIIE_weightedbyemployees), 
         alpha = type, 
         size = type,
         colour = firmsize,
         shape  = firmsize
         )) +
  geom_point(position = position_dodge(width = 0.7)) +
  scale_size_manual(values = c(1.5,3)) +
  scale_color_brewer(palette = 'Dark2') +
  scale_shape_manual(values = c(25,22,24)) +
  scale_alpha_manual(values = c(0.3,1)) +
  geom_vline(xintercept = 0, colour = 'black', alpha = 0.5) +
  xlab('AIIE') +
  ylab('') +
  guides(size = F,alpha=F) 
  # theme(legend.title=element_blank())


```



## UK Census 2021 Occupation data linked to Felten et al AI Occupational Exposure (AIOE): Greater Manchester local authorities + core cities

Census 2021 data at MSOA level contains 104 occupation categories. These are linked to Felten et al's AI Occupation Exposure (AIOE) and a weighted AIOE average worked out again (weighted by person counts in each occupation).

(Census data is used as it provides the best current mix of coverage and granularity, though it is now older).

So - this shows something different to the previous plot: the occupational exposure of **people who live in these MSOAs** (rather than firms' exposure in the above plot).

Things to note from this one:

* While the overall spread is less far from the mean (the AIOE is also normalised), some local authorities are clearly much more exposed, on this measure. For example, Trafford has very few MSOAs less than the AIOE mean.
* Residents in Greater Manchester and core cities' MSOAs tend to be on the 'more exposed' side of zero [**I will need to check this in more detail to make sure it holds up; I'd have expected central Manchester to have more of a mix?**]
* With some differences, the overall average order of local authorities doesn't differ hugely from that found in firm level data.

```{r}
#| fig-height: 9


#Via bulk download
#Workplace counts at MSOA first
#105 categories of SOC2020, so that's 
#3 digit "minor group"
#https://www.hesa.ac.uk/collection/coding-manual-tools/sicsocdata/soc-2020
occ.msoa <- read_csv("../../local/data/wp016/WP016_msoa.csv")
# occ.msoa <- read_csv("local/data/wp016/WP016_msoa.csv")

#Some tweaks to MSOA occupation counts census data
occ.msoa <- occ.msoa %>% 
  filter(!qg("does not apply",`Occupation (current) (105 categories) Label`)) %>% 
  mutate(
    occupation_name = str_sub(`Occupation (current) (105 categories) Label`,5,-1),
    occupation_code = str_sub(`Occupation (current) (105 categories) Label`,1,3)
  ) %>% 
  select(-c(`Occupation (current) (105 categories) Code`,`Occupation (current) (105 categories) Label`))


#Add AIOE into the MSOA data
occ.msoa.aioe <- occ.msoa %>% 
  rename(
    msoacode = `Middle layer Super Output Areas Code`,
    msoalabel = `Middle layer Super Output Areas Label`
    ) %>% 
  left_join(
    soc_3dig_AIOE %>% select(-SOC20203dig_name),
    by = c('occupation_code' = 'SOC20203dig_code')
  )



#From https://github.com/DanOlner/RegionalEconomicTools/blob/46907c9eb05193fe790579f73ff5f1ff019d90f5/quarto_docs/Bradford_sectorclusters.qmd#L384C1-L411C53
#England and Wales only here
corecities <- getdistinct('sheffield|Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Nottingham', unique(str_sub(occ.msoa.aioe$msoalabel,1,-5)))

#Yes, all in there, need to remove a few...
corecities <- corecities[!grepl(x = corecities, pattern = 'Greater|shire|side', ignore.case = T)]
corecities <- corecities[order(corecities)]

gmlas <- readRDS('../../data/gmlas.rds')
# gmlas <- readRDS('data/gmlas.rds')


#Per MSOA weighted means and labels
msoa.aioe.avs <- occ.msoa.aioe %>% 
  group_by(msoacode,msoalabel) %>% 
  summarise(
    AIOE_weightedmean = weighted.mean(AIOE,Count),
    sd_AIOE = sqrt(Hmisc::wtd.var(AIOE,Count))
  ) %>% 
  mutate(
    laname = str_sub(msoalabel,1,-5),
    type = case_when(
      laname %in% gmlas ~ "GM LA",
      laname %in% corecities[!qg('manc',corecities)] ~ "core city (minus manc)",
      .default = "other"
    ),
    is_GM = laname %in% gmlas
  )



#Plot individual MSOA values for each LA
ggplot(msoa.aioe.avs %>% filter(type != 'other'), 
       aes(
         x = AIOE_weightedmean, 
         y = fct_reorder(laname,AIOE_weightedmean), 
         colour = type, 
         size = is_GM)) +
  geom_jitter(height = 0.1) +
  # geom_errorbar(aes(xmin = xmin, xmax = xmax), width = 0.1) +
  scale_size_manual(values = c(0.5,1)) +
  # scale_colour_manual(values = c('blue','green','black')) +
  scale_color_brewer(palette = 'Dark2') +
  geom_vline(xintercept = 0, colour = 'black', alpha = 0.5) +
  xlab('AIOE') +
  ylab('') +
  guides(size = F) +
  theme(legend.title=element_blank())




```



# Initial results for 'jobs/tasks augmentable' versus 'jobs/tasks replaceable'

This section contains initial outputs from repeat LLM runs to extract probabilities for UK sectors having jobs/tasks that are (a) likely augmentable by AI; (b) likely replaceable by AI (running on a local instance of Llama 3.1).


## AIIE vs augmentability

This first plot does the following:

* X axis: Felten's AI Industrial Exposure (AIIE). Higher values are more exposed 1000m hexes in Greater Manchester (as in the interactive maps above).
* Y axis: "probability a job is more augmentable than replaceable (aug > replace)" within a sector (equivalent to complementarity index), applied to firms in Companies House.
* Circle size is employee count per 1000m hex.
* Broken down by Greater Manchester borough.

Some things to note:

* The same positive relationship seen in other results - more exposed jobs are on average also in more augmentable-jobs sectors
* Bottom right quadrant is the "high exposure plus jobs more likely to be replaceable" hexes. There aren't many of these, though most boroughs have one or two (with central Manchester having the most jobs in this quadrant). Note though, this is a probability - so e.g. an augmentable value of 0.6 would still on average suggest 40% of jobs in that sector could be replaceable.


```{r}
#| fig-height: 13

#Data for all firms with 1+ employees 
hexsummary <- readRDS('../../local/data/hexsummary.rds')

ggplot(
  hexsummary %>% rename(AIIE = AIIE_weightedbyemployees, `aug > replace` = augmentmorethanreplaceprob_weightedbyemployees),
  aes(x = AIIE, y = `aug > replace`, size = totalemployees)) +
  geom_point(alpha = 0.5) +
  scale_size(range = c(2,20)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0.5) +
  # geom_smooth(method = 'lm') +
  facet_wrap(~la,scales='free', ncol = 2)
  # facet_wrap(~la)


```


## Bivariate hexmaps for AIIE vs "aug > replace"

This and the next few plots are **bivariate maps** - each axis has three values of 'low / medium / high', making **nine possible combinations** of the two of them.

These two maps:

* Puts 'probability job is more augmentable than replaceable' **(aug>repl)** along one axis, with Felten's **AIIE** exposure index on the other. 
* The first is for **firms with 1 to 3 employees**, the second for **firms with 4+ employees**.
* Data for firms is summarised per 1000m hex (averages weighted by employee count and across any of the , as above).
* Hexes are only shown if they have a **min of 10 employees in them**.

So for example, in the map of firms with 1 to 3 employees:

* The darkest hexes in the south of Greater Manchester are high aug>rep (highest probability of jobs being more augmentable than replaceable) and also high AIIE (they have most AI exposure).
* There are only a few hexes with the legend's top left clear blue - high exposure but jobs more likely to be replaceable.
* In both maps, there are large areas of high exposure with the middle aug>rep probability (jobs likely a mix of both augmentable and replaceable).


### 1 to 3 employees

```{r}

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  # filter(Employees_thisyear > 9) %>% 
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= 10) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_polygons(
    fill = tm_vars(c("AIIE", "aug>rep"),
    # fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) 


```

### 4+ employees

```{r}

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  # filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  filter(Employees_thisyear > 4) %>%
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= 10) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_polygons(
    fill = tm_vars(c("AIIE", "aug>rep"),
    # fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) 


```




# 'Augmentable' and 'replaceable' each on their own scale

Here, two *separate* indices have been generated that each rank sectors as 'jobs more/less likely to be (1) replaceable and (2) augmentable in this sector compared to other sectors'.

These two indices are made by 'competing' repeated sectors against each other (using the probabilities assigned from the LLM) and putting these pairwise outcomes into a [Bradley Terry](https://en.wikipedia.org/wiki/Bradley%E2%80%93Terry_model){target="_blank"} model to produce a 'sector x more likely to be augmentable than y' ranking. These are then mean-centred (keeping the relative log-likelihoods between sectors intact).



## Where (270 3 digit SIC) sectors are located on the separate aug / replace axes

This first interactive plot just shows those two scales on each axis to illustrate where the method places each sector - **hover for sector name**. 


```{r}

bt_trials_wsectors <- readRDS('../../local/data/bt_trials.rds')
#bt_trials_wsectors <- readRDS('local/data/bt_trials.rds')

p <- ggplot(bt_trials_wsectors,aes(x = replace, y = augment, label = SIC_3DIGIT_NAME)) +
  geom_point(size = 3) +
  geom_vline(xintercept = 0, colour = 'darkgrey') +
  geom_hline(yintercept = 0, colour = 'darkgrey') 

ggplotly(p, tooltip = 'label')




```


## Separate augment / replace hexmaps

These maps put 'probability augmentable (high/medium/low)' and 'probability replaceable (high/medium/low)' on separate axes, again meaning there are nine possible values.

As above, there is one map for 'firms with 1 to 3 employees' and a second for 'firms with 4+ employees'.

Some comments on these two maps:

* There's an obvious stark difference for 1-3 employee vs 4+ employee firms - for 1-3s, there are large areas with lower augmentable and medium to high replacable.
* Though there's a cluster (matching the one above for aug>repl vs AIIE) that's low replaceable / medium augmentable (which fits with 'prob aug is more than prob replace' but here we can get a better idea of the actual scale of both).
* The 4+ map has a lot more hexes in the max category for *both* aug and replace.


### Firms with 1 to 3 employees

```{r}

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  # filter(Employees_thisyear > 4) %>%
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= 10) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_polygons(
    # fill = tm_vars(c("AIIE", "aug>rep"),
    fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) 


```



### Firms with 4+ employees

```{r}

hexsummary <- overlay %>% 
  st_set_geometry(NULL) %>% 
  # filter(Employees_thisyear > 0) %>% #Only firms with employees recorded in latest year
  # filter(between(Employees_thisyear,1,3)) %>% #Microfirms
  # filter(between(Employees_thisyear,4,9)) %>% #Microfirms
  filter(Employees_thisyear > 4) %>%
  group_by(id) %>% 
  summarise(
    AIIE_weightedbyemployees = weighted.mean(AIIEfinal,Employees_thisyear),
    augmentmorethanreplaceprob_weightedbyemployees = weighted.mean(augment_morethan_replace_prob,Employees_thisyear),
    augmentBT_weightedbyemployees = weighted.mean(augment_logability_BT,Employees_thisyear),
    replaceBT_weightedbyemployees = weighted.mean(replace_logability_BT,Employees_thisyear),
    totalemployees = sum(Employees_thisyear),
    totalfirms = n(),
    la = max(localauthority_name)
  ) %>% 
  group_by(id) %>%
  filter(sum(totalemployees) >= 10) %>% #keep only gridsquares where total employee count is more than / equal to 100
  ungroup()

#Link that back into the grid squares...
#Use right join to drop empties
sq.ch <- sq %>% 
  right_join(
    hexsummary,
    by = 'id'
  )

tm_shape(sq.ch %>% rename(aug = augmentBT_weightedbyemployees, repl = replaceBT_weightedbyemployees, AIIE = AIIE_weightedbyemployees, `aug>rep` = augmentmorethanreplaceprob_weightedbyemployees)) +
  tm_polygons(
    # fill = tm_vars(c("AIIE", "aug>rep"),
    fill = tm_vars(c("aug", "repl"),
                   multivariate = TRUE),
              fill.scale = 
                tm_scale_bivariate(
                  scale1 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  scale2 = tm_scale_intervals(style = "fisher", n = 3, labels = c("L", "M", "H")),
                  # values = "stevens.bluered")) +
                  values = "bu_br_bivs")) +
  tm_shape(lad) +
  tm_borders(col = 'black', lwd = 1, fill_alpha = 0.3) +
  tm_shape(itl2) +
  tm_borders(col = 'black', lwd = 4) 


```








# Extra bits

## Same AIOE occupation exposure data, but just at local authority level

I've included this just to make a quick point. This is the same data as the last plot, but with weighted mean and standard deviation applied at the local authority level. 

The points are the means; bars are one standard deviation. 

While the means do differ, each local authority's spread of AIOE covers a very similar level of exposure from top to bottom. **What I'm taking from this: important differences show up at the sub-local authority level.**

```{r}
#| fig-height: 9

#Upper tier
occ.la <- read_csv("../../local/data/wp016/WP016_utla.csv")
# occ.la <- read_csv("local/data/wp016/WP016_utla.csv")

#Some tweaks to local authority occupation counts census data
occ.la <- occ.la %>% 
  filter(!qg("does not apply",`Occupation (current) (105 categories) Label`)) %>% 
  mutate(
    occupation_name = str_sub(`Occupation (current) (105 categories) Label`,5,-1),
    occupation_code = str_sub(`Occupation (current) (105 categories) Label`,1,3)
  ) %>% 
  select(-c(`Occupation (current) (105 categories) Code`,`Occupation (current) (105 categories) Label`))


#Add AIOE into the Census data
occ.la.aioe <- occ.la %>% 
  left_join(
    soc_3dig_AIOE %>% select(-SOC20203dig_name),
    by = c('occupation_code' = 'SOC20203dig_code')
  )

#Check per LA means
la.aioe.avs <- occ.la.aioe %>% 
  group_by(`Upper tier local authorities Label`) %>% 
  summarise(
    AIOE_weightedmean = weighted.mean(AIOE,Count),
    sd_AIOE = sqrt(Hmisc::wtd.var(AIOE,Count))
    )

la.aioe.avs <- la.aioe.avs %>%
  rename(laname = `Upper tier local authorities Label`) %>% 
  mutate(type = case_when(
    laname %in% gmlas ~ "GM LA",
    laname %in% corecities[!qg('manc',corecities)] ~ "core city (minus manc)",
    .default = "other"
  ),
  is_GM = laname %in% gmlas,
  xmin = AIOE_weightedmean - (sd_AIOE),
  xmax = AIOE_weightedmean + (sd_AIOE)
  )


ggplot(la.aioe.avs %>% filter(type != 'other'), 
       aes(
         x = AIOE_weightedmean, 
         y = fct_reorder(laname,AIOE_weightedmean), 
         colour = type, 
         size = is_GM)) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = xmin, xmax = xmax), width = 0.1) +
  scale_size_manual(values = c(0.5,1)) +
  # scale_colour_manual(values = c('blue','green','black')) +
  scale_color_brewer(palette = 'Dark2') +
  geom_vline(xintercept = 0, colour = 'black', alpha = 0.5) +
  xlab('AIOE') +
  ylab('') +
  guides(size = F) +
  theme(legend.title=element_blank())



```
